Evaluating Cloud-Optimized HDF5 for NASA’s ICESat-2 Mission

### 2025-01-15 paper.qmd
efficient access will accelerate science

abstract: |
  The Hierarchical Data Format (HDF) is a common archival format for n-dimensional scientific data; it has been utilized to store valuable information from astrophysics to earth sciences and everything in between. As flexible and powerful as HDF can be, it comes with big tradeoffs when files on remote storage systems are accessed using HTTPS (Hypertext Transfer Protocol Secure), mainly because the file format and the client I/O libraries were designed for local and supercomputing workflows. As scientific data and workflows migrate to the cloud, efficient access to data stored in HDF format is a key factor that will accelerate or slow down “science in the cloud” across all disciplines.
  We present an implementation of recently available features in the HDF5 stack that results in performant access to HDF from remote cloud storage. This performance is on par with modern cloud-native formats like Zarr but with the advantage of not having to reformat data or generate metadata sidecar files (DMR++, Kerchunk). Our benchmarks also show potential cost-savings for  data producers if their data are processed using cloud-optimized strategies.

>>>>

abstract: |
  The Hierarchical Data Format (HDF) is a common archival format for n-dimensional scientific data; it has been utilized to store valuable information from astrophysics to earth sciences and everything in between. As flexible and powerful as HDF can be, it comes with big tradeoffs when files on remote storage systems are accessed using HTTPS (Hypertext Transfer Protocol Secure), mainly because the file format and the client I/O libraries were designed for local and supercomputing workflows. As scientific data and workflows migrate to the cloud, efficient access to data stored in HDF format is a key factor that will accelerate “science in the cloud” across all disciplines.
  We present an implementation of recently available features in the HDF5 stack that results in performant access to HDF from remote cloud storage. This performance is on par with modern cloud-native formats like Zarr but with the advantage of not having to reformat data or generate metadata sidecar files (DMR++, Kerchunk). Our benchmarks also show potential cost-savings for  data producers if their data are processed using cloud-optimized strategies.
